Concurrent Programming in Harmony

### 2021-08-09 book/paper.tex
Latest

The Needham-Schroeder protocol~\cite{NS78} is a security protocol in which two parties
authenticate one another by exchanging large and recently created random numbers
called \emph{nonces} that nobody else should be able to read.
The version we describe here uses \emph{public key cryptography}~\cite{DH76}: with public key
cryptography it is possible to create a message for a particular destination that
only that destination can read.  We denote with $\langle m \rangle_p$ a message
$m$ encrypted for $p$ so that only $p$ can decrypt the message and see that it contains $m$.

@@

When Bob receives $N_B$ in the last step, it decides that it is Alice that is
trying to communicate at this time.  Since Bob created $N_B$ recently and sent it
encrypted to Alice, Bob does not have to worry that the message was an old message
that was replayed.  Also, if Alice is honest, only Alice can have seen the message
containing $N_B$.

@@

\begin{enumerate}
\item Alice creates a new nonce $N_A$ and sends $\langle 1, A, N_A \rangle_\mathtt{Corey}$
to Corey;
\item Upon receipt, Corey sends $\langle 1, A, N_A \rangle_\mathtt{Bob}$ to Bob;
\item Upon receipt, Bob creates a new nonce $N_B$ and sends
$\langle 2, N_A, N_B \rangle_\mathtt{Alice}$ to Alice;
\item Alice thinks the message came from Corey (because it contains $N_A$, which
Alice sent to Corey) and
sends $\langle 3, N_B \rangle_\mathtt{Corey}$ to Corey.
\item Corey learns $N_B$, which it was not supposed to.
\end{enumerate}

Corey can use $N_B$ to send $\langle 3, N_B \rangle_\mathtt{Bob}$ to Bob.
Bob now thinks it is talking to Alice, even though Alice never tried to
communicate with Bob.

>>>>

The Needham-Schroeder protocol~\cite{NS78} is a security protocol in which two parties
authenticate one another by exchanging large and recently created random numbers
called \emph{nonces} that nobody else should be able to read.
The nonces should only be used once for an instantiation of the protocol between
honest participants.
The version we describe here uses \emph{public key cryptography}~\cite{DH76}: with public key
cryptography it is possible to create a message for a particular destination that
only that destination can read.  We denote with $\langle m \rangle_p$ a message
$m$ encrypted for $p$ so that only $p$ can decrypt the message and see that it contains $m$.

@@

When Bob receives $\langle 1, A, N_A \rangle_\mathtt{Bob}$, Bob does not know for sure
that the message came from Alice, and even if it came from Alice, it does not know if
Alice sent the message recently or if it was replayed by some adversary.
When Alice receives $\langle 2, N_A, N_B \rangle_\mathtt{Alice}$, Alice \emph{does}
know that, if Bob is honest, (1) Bob and only Bob could have created this message, and
(2) Bob must have done so recently (since Alice created $N_A$).
When Bob receives $\langle 3, N_B \rangle_\mathtt{Bob}$, Bob decides that it is Alice
that is trying to communicate at this time.  Since Bob created $N_B$ recently and sent it
encrypted to Alice, Bob does not have to worry that the type 3 message was an old message
that was replayed by some adversary.
Also, if Alice is honest, only Alice can have seen the message containing $N_B$.

Thus, the intended security property of this protocol is that if Alice and Bob are both
honest, only Alice and Bob learn $N_A$ and $N_B$ and only if they are trying to communicate
with one another.

@@

\begin{enumerate}
\item Alice creates a new nonce $N_A$ and sends $\langle 1, A, N_A \rangle_\mathtt{Corey}$
to Corey;
\item Upon receipt, Corey sends $\langle 1, A, N_A \rangle_\mathtt{Bob}$ to Bob;
\item Upon receipt, Bob, as usual, creates a new nonce $N_B$ and sends
$\langle 2, N_A, N_B \rangle_\mathtt{Alice}$ to Alice;
\item Alice thinks the message came from Corey (because it contains $N_A$, which
Alice created for Corey and sent to Corey) and
sends $\langle 3, N_B \rangle_\mathtt{Corey}$ to Corey.
\item Corey learns $N_B$, which it was not supposed to as Bob is only willing to
talk to Alice and created $N_B$ explicitly for Alice.
\end{enumerate}

Corey can use $N_B$ to send $\langle 3, N_B \rangle_\mathtt{Bob}$ to Bob.
This is identical to the situation in which Alice tried to set up a connection to Bob,
so Bob now thinks it is talking to Alice, even though Alice never tried to
communicate with Bob.  The security property is violated because Corey discovers $N_B$
even though Alice and Bob are both honest and are not trying to communicate with one
another.  As a result, Carter has successfully impersonated Alice to Bob.
